### 2019-05-10 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Using \begin{framed}...\end{framed} for all algorithms.
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

We assume to have $N \geq 2$ MAB algorithms, $\Alg_1, \dots, \Alg_N$,
and let $\Alg_{\mathrm{aggr}}$ be an aggregation algorithm,
which runs the $N$ algorithms in parallel (with the same slotted time), and use them to choose its channels based on a voting from their $N$ decisions.
%
$\Alg_{\mathrm{aggr}}$ depends on a pool of algorithms and a set of parameters.
We would like that $\Alg_{\mathrm{aggr}}$
performs almost as well as the best of the $\Alg_a$, with a good choice of its parameters, independently of the MAB problem.
Ideally $\Alg_{\mathrm{aggr}}$ should perform similarly to the best of the $\Alg_a$.
%
To simplify the presentation, we only aggregate bandit algorithms that give deterministic recommendations:
one arm is chosen with probability $1$ and the others with probability $0$.
However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $\{1,\dots,K\}$ at each time step, and draw the next selected arm according to this distribution.

The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$, starting from a uniform distribution:
$\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in \{1,\dots,N\}$ from $\pi^t$, then follows $\Alg_a$'s decision.
The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very points.

@@

Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
%
\begin{small}
	\begin{figure}[h!]
		\centering
		% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
		% Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
		\begin{algorithm}[H]
			% XXX Input, data and output
			\KwIn{$N$ bandit algorithms, $\Alg_1, \dots, \Alg_N$, with $N \geq 2$}
			\KwIn{Number of arms, $K \geq 2$}
			\KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
			\KwData{Initial uniform distribution, $\pi^{0} = U(\{1, \dots, N\})$}
			\KwResult{$\Alg_{\mathrm{aggr}}=\Aggr\left[\Alg_1, \dots, \Alg_N\right]$}
			% XXX Algorithm
			\For(\tcp*[f]{At every time step})
            {$t = 1, \dots, T$}{
				%
				% \tcp{First, run each algorithm $\Alg_a$}
				\For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
					$\Alg_a$ updates its indexes (e.g., \UCB{} indexes)\;
					It chooses $A^{t+1}_{a} \in \{1, \dots, K\}$.
				}
				%
				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A^{t+1}_{a} = j\}), \;\forall 1\leq j \leq K$\;
				% \tcp*{Sum trusts}
				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A^{t+1} \sim p^{t+1}$\;
				% $\P(A^{t+1}_{\mathrm{aggr}} = j) = p^{t+1}_j$
				% \tcp*{Sample once}
				%
				Give \emph{original} reward $(A^{t+1}, 1 - \ell_{A^{t+1}}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A^{t+1}}(t+1) / p^{t+1}_{A^{t+1}}\;. $$
				% \tcp{Use bandit feedback to update $\pi^{t}$}
				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A^{t+1},t}$) to update $\pi^{t}$}
				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
				\For{$a = 1, \dots, N$}{
					% \eIf{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
					\If{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{More trusted}
						}%{
						% $ \pi^{t+1}_{a} = \exp(- \eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{Less trusted}
					% }
				}
				Renormalize the new $\pi$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
				% \tcp*{Project it back to $\Delta_N$}
			}
			\caption{Our aggregation algorithm \Aggr.}
			\label{algo:25:Aggr}
		\end{algorithm}
	\end{figure}
\end{small}

>>>>

We assume to have $N \geq 2$ MAB algorithms, $\Alg_1, \dots, \Alg_N$,
and let $\Alg_{\mathrm{aggr}}$ be an aggregation algorithm,
which runs the $N$ algorithms in parallel (with the same slotted time), and use them to choose its channels based on a voting from their $N$ decisions.
%
$\Alg_{\mathrm{aggr}}$ depends on a pool of algorithms and a set of parameters.
We would like that $\Alg_{\mathrm{aggr}}$
performs almost as well as the best of the $\Alg_a$, with a good choice of its parameters, independently of the MAB problem.
Ideally $\Alg_{\mathrm{aggr}}$ should perform similarly to the best of the $\Alg_a$.
%
To simplify the presentation, we only aggregate bandit algorithms that give deterministic recommendations:
one arm is chosen with probability $1$ and the others with probability $0$.
However, both \ExpQ{} and \Aggr{} can be adapted to aggregate randomized bandit algorithms, \ie, algorithms that output a probability distribution $\xi_t$ over the arms $[K]$ at each time step, and draw the next selected arm according to this distribution.

The aggregation algorithm maintains a probability distribution $\pi^{t}$ on the $N$ algorithms $\Alg_a$, starting from a uniform distribution:
$\pi^t_a$ is the probability of trusting the decision made by algorithm $\Alg_a$ at time $t$.
$\Alg_{\mathrm{aggr}}$ then simply performs a weighted vote on its algorithms: it decides whom to trust by sampling $a \in [N]$ from $\pi^t$, then follows $\Alg_a$'s decision.
The main questions are then to know what observations (\ie, arms and rewards) should be given as feedback to which algorithms,
and how to update the trusts at each step, and our proposal \Aggr{} differs from \ExpQ{} on these very points.

@@

Our proposed \Aggr{} is detailed in Algorithm~\ref{algo:25:Aggr}.
%
\begin{small}
	\begin{figure}[h!]
		\centering
		\begin{framed}
		% Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
		% Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
		\begin{algorithm}[H]
			% XXX Input, data and output
			\KwIn{$N$ bandit algorithms, $\Alg_1, \dots, \Alg_N$, with $N \geq 2$}
			\KwIn{Number of arms, $K \geq 2$}
			\KwIn{Time horizon, $T \geq 1$, \textbf{not} used for the learning}
			\KwIn{A sequence of learning rates, $(\eta_t)_{t \geq 1}$}
			\KwData{Initial uniform distribution, $\pi^{0} = U([N])$}
			\KwResult{$\Alg_{\mathrm{aggr}}=\Aggr\left[\Alg_1, \dots, \Alg_N\right]$}
			% XXX Algorithm
			\For(\tcp*[f]{At every time step})
            {$t = 1, \dots, T$}{
				%
				% \tcp{First, run each algorithm $\Alg_a$}
				\For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
					$\Alg_a$ updates its indexes (e.g., \UCB{} indexes)\;
					It chooses $A^{t+1}_{a} \in [K]$.
				}
				%
				% \tcp{The more algorithms have chosen arm $j$, the more probable it is to be chosen.}
				Let $p^{t+1}_j := \sum\limits_{a = 1}^{N} \pi^{t}_a \times \mathbbm{1}(\{A^{t+1}_{a} = j\}), \;\forall 1\leq j \leq K$\;
				% \tcp*{Sum trusts}
				Then $\Alg_{\mathrm{aggr}}$ chooses arm $A^{t+1} \sim p^{t+1}$\;
				% $\P(A^{t+1}_{\mathrm{aggr}} = j) = p^{t+1}_j$
				% \tcp*{Sample once}
				%
				Give \emph{original} reward $(A^{t+1}, 1 - \ell_{A^{t+1}}(t+1))$ to \emph{each} $\Alg_a$ (maybe \emph{not} on its chosen arm)\;
				Compute an \emph{unbiased} estimate of the loss of the trusted algorithms,$$ \ell^{t+1} = \ell_{A^{t+1}}(t+1) / p^{t+1}_{A^{t+1}}\;. $$
				% \tcp{Use bandit feedback to update $\pi^{t}$}
				% \tcp{Finally, use the instantaneous bandit feedback (loss $\ell_{A^{t+1},t}$) to update $\pi^{t}$}
				% \For(\tcp*[f]{Can be parallel}){$a = 1, \dots, N$}{
				\For{$a = 1, \dots, N$}{
					% \eIf{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
					\If{$\Alg_a$ was trusted, \ie, $A^{t+1}_{a} = A^{t+1}$}{
						$ \pi^{t+1}_{a} = \exp(\eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{More trusted}
						}%{
						% $ \pi^{t+1}_{a} = \exp(- \eta_t \ell^{t+1}) \times \pi^{t}_{a} $
						% \tcp*{Less trusted}
					% }
				}
				Renormalize the new $\pi$: $\pi^{t+1} := \pi^{t+1} / \sum_{a=1}^{N} \pi^{t+1}_{a}$.
				% \tcp*{Project it back to $\Delta_N$}
			}
			\caption{Our aggregation algorithm \Aggr.}
			\label{algo:25:Aggr}
		\end{algorithm}
		\end{framed}
	\end{figure}
\end{small}

### 2019-05-10 2-Chapters/3-Chapter/chapter3.tex
...
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
a stochastic MAB problem is defined by $K>1$ distributions $\nu_k$ (also called arms),
used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
An agent chooses arm $A(t)\in\{1,\dots,K\}$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.

@@

In this section, we report additional experiment results from the simulations described in the previous Section~\ref{sec:3:reviewSPAlgorithms}.
Instead of studying on the efficiency of the algorithms (\ie, their regret), we report results of real measurements in terms of computational time as well as memory storage.
%
The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in\{1,\dots,T\}$ bounded by $\bigO{K}$, independent of $t$.

@@

\pagagraph{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
\begin{itemize}
    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in\{1,\dots,T\}$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
    This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
    On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
\end{itemize}
In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
%     def time()
% time() -> floating point number

>>>>

For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
a stochastic MAB problem is defined by $K>1$ distributions $\nu_k$ (also called arms),
used to sample \iid{} rewards $r_k(t) \sim \nu_k, \forall t$.
An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r_{A(t)}(t)$ without knowing the other (hidden) rewards.
Her goal is to maximize $\sum_{t=1}^T r_{A(t)}(t)$ by sequentially exploring the $K$ arms, and she essentially has to find and exploit the best one as fast as possible.

@@

In this section, we report additional experiment results from the simulations described in the previous Section~\ref{sec:3:reviewSPAlgorithms}.
Instead of studying on the efficiency of the algorithms (\ie, their regret), we report results of real measurements in terms of computational time as well as memory storage.
%
The goal of this section is to verify that the simplest (but most efficient) algorithms have a memory cost proportional to the number of arms but independent of the horizons, \ie, bounded by $\bigO{K}$, and a time complexity at each time step $t\in[T]$ bounded by $\bigO{K}$, independent of $t$.

@@

\paragraph{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.
\begin{itemize}
    \item For time, we used the \texttt{time.time()} function from Python's standard library, that gives the current time in micro-seconds.
    In the \texttt{Evaluator} object of SMPyBandits, before launching the ``for'' loop on $t\in[T]$ for one algorithm, the system time is stored, and when the loop is finished, the time used for this loop is the current system time minus the starting time.
    This measurement gets averaged on the $N=100$ (independent) repetitions, and consistently measure the time efficiency of the algorithms.
    \item For memory, we use two different approaches whether the code runs on a UNIX or a Windows system.
    On UNIX, the \texttt{resource} module allows to measure the memory of the current process, and by counting the difference in memory used by the \texttt{Evaluator} object before and after the ``for'' loop, we can track the memory used by the algorithm to store all its interior variables.
\end{itemize}
In both cases, the ``for'' loop takes some time and stores many variables, but the only difference in terms of both time or memory between two algorithms is explained by the difference in the implementation of the algorithms.
%     def time()
% time() -> floating point number

### 2019-05-10 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Talk to Emilie
Done changing notations \{1,\dots,M\} -> [M] all over the PhD

\paragraph{MAB model}
%
We consider the network from the point of view of one dynamic device.
Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
which is a usual objective in MAB problems.

@@

Figure \ref{fig:41:from10to100} presents the evolution of the successful transmission rate, as a function of time.
The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.

>>>>

\paragraph{MAB model}
%
We consider the network from the point of view of one dynamic device.
Following a Bernoulli emission process, every times a dynamic device has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in[K]$.
Then, the dynamic device waits in this channel $C(t)$ for an acknowledgement sent by the gateway, during a certain period (\eg, $5$ seconds for the example considered above).
Before sending another message (\ie, at time $t+1$), the dynamic device knows if it received or not this \Ack{} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a (binary) \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \Ack{} was received before the next message, or $1$ if \Ack{} was successfully received.
The goal of the dynamic device is to minimize its packet loss ratio, or equivalently, to maximize its number of successful transmissions, or its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
which is a usual objective in MAB problems.

@@

Figure \ref{fig:41:from10to100} presents the successful transmission rate as a function of time.
The two MAB algorithms, \UCB{} and Thompson Sampling (TS), are compared against the naive random policy (which are outperformed easily by the MAB algorithms), and the two (optimal and greedy) oracle policies (which outperform slightly the MAB algorithms).
The results are displayed when $10\%$, $30\%$, $50\%$ and $100\%$ of the traffic is generated by dynamic devices.
